

<center>
# STA 202 Project 1 Group 3<br>
Kshitij Shekhar-AU2020165<br>
Laksh Gupta-AU2020152<br>
Vinayak C Manoj-AU2020015<br>
</center>
<br>

## Question 1
(A)**Suppose $f(x)=\frac{n!}{x!(n-x)!}p^x(1-p)^{n-x}$, where X can take values from 1 to n
Take, n = 20 and p = 0.5. Write some R code to plot the probability mass function and and
check if the function is a probability probability function or not?**
</center>
```{r}
x = 20
n = 20
p = 0.5

fx = array(0, dim=x) # array of size 20 to store the values of f(x)
X = array(0, dim=x) # array of size 20 to store the values of corresponding x

for (i in 0:x)
{
  f = (factorial(n)/(factorial(i)*factorial(n-i)))*(p^i)*((1-p)^(n-i)) # f(x) by definition
  X[i] = i # appending the values of x in 'X'
  fx[i] = f # appending the values of f(x) in 'fx'
}

plot(X, fx, xlab = "x", ylab = "f(x)", main = "Probability Mass Function", type='h') # plotting the probability mass function
```
<br>As $\sum_{x=0}^n f(x)=1$, as shown below, this is a valid probability mass function.
```{r}
x = 20
n = 20
p = 0.5
g = 0 # defined to add values of each f(x)

for (i in 0:x)
{
  f = (factorial(n)/(factorial(i)*factorial(n-i)))*(p^i)*((1-p)^(n-i))
  g = g + f # sum of all f(x)
  #print(f)
}

print (g)
```


(B)**Mention the name of the probability distribution.**

This distribution is called the Binomial Distribution

(C)**Write R codes to draw a random sample of size n = 20, and find sample mean and
sample variance of the probability distribution X.**
```{r}
x = 20
n = 20
p = 0.5

fx = array(0, dim=x) # array of size 20 to store the values of f(x)

for (i in 0:x)
{
  f = (factorial(n)/(factorial(i)*factorial(n-i)))*(p^i)*((1-p)^(n-i)) # f(x) by definition
  fx[i] = f # appending the values of f(x) in 'fx'
}

y = sample(fx, 20, replace = TRUE) # sampling data from f(x) with allowed replacement
print(y)

sample_mean = mean(y) #sample mean
sample_variance = var(y) #sample variance

print(sample_mean)
print(sample_variance)
```

## Question 2

```{r}
f <- function(x) {
  (1/2*sqrt(2*pi)) * exp(-0.125 * (x-2)^2)
}

# Plot the function
curve(f, from = -10, to = 10)

# Check condition 1: f(x) is non-negative
all(sapply(seq(-10, 10, by = 0.01), f) >= 0) # should return TRUE

# Check condition 2: the area under the curve is equal to 1
integrate(f, lower = -10, upper = 10)

# Set the seed for reproducibility
set.seed(12)

# Values to sample
x_vals <- seq(-5, 10, 0.015)

# Apple PDF to x_vals
y_vals <- sapply(x_vals, f)

samples <- sample(x_vals, 50, replace = TRUE, prob = y_vals)

# Find the sample mean and standard error
sample_mean <- mean(samples)
sample_se <- sd(samples) / sqrt(length(samples))

# Print the results
print(paste("Sample mean:", sample_mean))
print(paste("Sample standard error:", sample_se))
```


## Question 3

(A) **Consider the data given in the above example of Chinese population in Hong Kong in
1937 and estimate the maximum likelihood estimate (MLE) of the parameter θ as $\hat{\theta}$ and
estimate the three gen0type probabilities for M, MN and N.**
<center>
Let $X_1, X_2. X_3$ denote the counts in the three cells. $n=1029$. We can find the log likelihood of $\theta$
as follows:
<center><br>
$l(\theta)= \log(n!) - \sum_{i=1}^3 log X_i! + X_1log(1-\theta)^2 + X_2log2\theta(1-\theta) + X_3log\theta^2$
<center><br>
$=\log(n!) - \sum_{i=1}^3 log X_i! + (2X_1 + X_2)log(1-\theta) + (2X_3 + X_2)log\theta + X_2log2$

Setting the derivative to zero to maximize $l(\theta)$,
<center><br>
$-\frac{2X_1+X_2}{1-\theta} + \frac{2X_3 + X_2}{\theta} = 0$,
</center>
On solving the above equation, we get $\hat{\theta} = 0.4247$

According to the Hardy-Weinberg Equilibrium, if the blood types are in equilibrium, alleles M, MN and N occur in a population with probabilities $(1-\theta)^2, 2\theta(1-\theta),$ and $\theta^2$ respectively.
Since we do not know the actual value of $\theta$, we can use $\hat{\theta}$ according to the bootstrap principle.<br>
$\therefore$ (M,MN,NN) occur with probability $.331,.489,.180$ 

(B)**How do we assess the preciseness of the estimate $\hat{\theta}$? Make comments.**

We can assess the preciseness of $\hat{\theta}$ by using the bootstrap method to estimate the sampling distribution with a histogram, and by extension its standard error


(C)**Estimate the sampling distribution of θ (ie, draw the histogram) and report the standard
error.** 
</center>
```{r}

observed_freqs=c(342,500,187)
# Calculate the observed value of theta
f_M <- observed_freqs[1]
f_MN <- observed_freqs[2]
f_N <- observed_freqs[3]


p_hat <- (2 * f_M + f_MN) / (2 * (f_M + f_MN + f_N))
q_hat <- (2 * f_N + f_MN) / (2 * (f_M + f_MN + f_N))
theta_hat <- 2 * p_hat * q_hat

# Set the number of bootstrap samples
n_bootstraps <- 1000

# Create an empty vector to store the bootstrap estimates of theta
theta_hat_boot <- numeric(n_bootstraps)

# Run the bootstrap
for (i in 1:n_bootstraps) {
  # Sample with replacement from the observed frequencies
  boot_sample <- sample(observed_freqs, size = length(observed_freqs), replace = TRUE)
  
  # Calculate the estimated value of theta in the bootstrap sample
  f_M_boot <- boot_sample[1]
  f_MN_boot <- boot_sample[2]
  f_N_boot <- boot_sample[3]
  p_hat_boot <- (2 * f_M_boot + f_MN_boot) / (2 * (f_M_boot + f_MN_boot + f_N_boot))
  q_hat_boot <- (2 * f_N_boot + f_MN_boot) / (2 * (f_M_boot + f_MN_boot + f_N_boot))
  theta_hat_boot[i] <- 2 * p_hat_boot * q_hat_boot
}

# Calculate the standard error of the bootstrap estimates of theta
se_theta_hat_boot <- sd(theta_hat_boot)

# Print the estimated value of theta and the standard error
cat("Estimated value of theta:", theta_hat, "\n")
cat("Standard error of theta:", se_theta_hat_boot, "\n")
CI <- quantile(theta_hat_boot, c(0.05, 0.95))
print(CI)
hist(theta_hat_boot)
```

(D)**Use bootstrap method (take number of bootstrap sample 500) to estimate the sampling
distribution and standard error of the ˆθ and 90% confidence interval of θ.**

```{r}
n_bootstraps <- 500

# Create an empty vector to store the bootstrap estimates of theta
theta_hat_boot <- numeric(n_bootstraps)

# Run the bootstrap
for (i in 1:n_bootstraps) {
  # Sample with replacement from the observed frequencies
  boot_sample <- sample(observed_freqs, size = length(observed_freqs), replace = TRUE)
  
  # Calculate the estimated frequencies of each blood type in the bootstrap sample
  f_M_boot <- boot_sample[1]
  f_MN_boot <- boot_sample[2]
  f_N_boot <- boot_sample[3]
  
  # Calculate the estimated allele frequencies in the bootstrap sample
  p_hat_boot <- (2 * f_M_boot + f_MN_boot) / (2 * (f_M_boot + f_MN_boot + f_N_boot))
  q_hat_boot <- (2 * f_N_boot + f_MN_boot) / (2 * (f_M_boot + f_MN_boot + f_N_boot))
  
  # Calculate the estimated theta in the bootstrap sample
  theta_hat_boot[i] <- 2 * p_hat_boot * q_hat_boot
}

# Calculate the standard error of the bootstrap estimates of theta
se_theta_hat_boot <- sd(theta_hat_boot)

# Calculate the 90% confidence interval for theta
CI <- quantile(theta_hat_boot, c(0.05, 0.95))
print(CI)
#print(c(theta_hat_boot - 1.645 * se_theta_hat_boot, theta_hat_boot + 1.645 * se_theta_hat_boot))
```


(E)**Use large sample method to estimate SE and the 90% CI for $\theta$**

We know that Standard Error is given by the following:
$SE=\sqrt{\hat{p}\hat{q}/n}$<br>
where $\hat{p}$ and $\hat{q}$ are the sample frequencies of the M and N alleles respectively, n is the sample size

The 90% confidence interval is calculated as follows:
$(\hat{\theta} - (1.645*SE), \hat{\theta} + (1.645*SE))$<br>
where 1.645 is the Z score corresponding to a 90% CI

```{r}
# Sample frequencies of M, MN, and N blood types
f_M <- 342 / 1029
f_MN <- 500 / 1029
f_N <- 187 / 1029

# Sample size
n <- 1029

# Estimate of theta
theta_hat <- f_MN

# Estimate of p and q
p_hat <- 2 * f_M + f_MN
q_hat <- 2 * f_N + f_MN

# Standard error
SE <- sqrt((p_hat * q_hat) / n)

# 90% confidence interval
CI <- c(theta_hat - 1.645 * SE, theta_hat + 1.645 * SE)

# Print results
cat("Standard Error:", SE, "\n")
cat("90% Confidence Interval:", CI, "\n")
```



## Contributions in the project
**Question 1** : Laksh Gupta <br>

**Question 2** : Vinayak Manoj<br>

**Question 3** : Kshitij Shekhar

**Report** : Kshitij Shekhar
